Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2014-09-19 paper/paper.tex
various tweaks to paper from reading it over Tested everything on bennet.

However, Mannik \emph{et al.} have also shown that there are
limitations to this robust capability to
oscillate~\cite{mannik2010bacteria,mannik2009bacterial}. They have
experimentally forced E. Coli cells into silicon microfabricated
channels of $0.25\mu m$ thickness. Upon entering the channels, the
cells undergo a rapid mechanical deformation in which they widen along
their short axis parallel to the channel plane, followed by a slower,
broadening out along both free axes.  This deformation results in very
wide (they can reach widths of over $5\mu m$), flattened cells that
when view from the top down show very irregular and assymetric shapes.
While these cells are still able to divide into surprisingly equal
volumes, the MinD oscillations within lose their regularity, both
temporally and spatially. Seen with flourescent microscopy (after
using common genetic techniques to allow the cells to flouresce), the
MinD seem to maximize in multiple locations within the cell, in a
seemingly random sequence. This allows for an opportunity to test
previously studied MinD simulation models against more extreme
experimental cases than have been seen thus far. \fixme{Perhaps this
  sentence should come later after we've introduced the simulation
  research that's been done.}

@@

\begin{multline}
  \frac{\partial \rho_{D:ADP}}{\partial t} = \mathcal{D}_D\nabla^2\rho_{D:ADP}-k_D^{ADP\rightarrow ATP}\rho_{D:ADP}\\
  +\delta(d_w)k_{de}\sigma_{DE},\hspace{3.4cm}
\end{multline}
\begin{multline}
  \frac{\partial \rho_{D:ATP}}{\partial t} = \mathcal{D}_D\nabla^2\rho_{D:ATP}+k_D^{ADP\rightarrow ATP}\rho_{D:ADP}\\
  -\delta(d_w)[k_D+k_{dD}(\sigma_D+\sigma_{DE})]\rho_{D:ATP}
\end{multline}
\begin{multline}
  \frac{\partial \rho_E}{\partial t} = \mathcal{D}_E\nabla^2\rho_E+\delta(d_w)k_{de}\sigma_{DE}
  -\delta(d_w)k_E \sigma_D \rho_E
\end{multline}
\begin{multline}
  \frac{\partial \sigma_D}{\partial t} = -k_E\sigma_D\rho_E
  +[k_D+k_{dD}(\sigma_D+\sigma_{DE})]\rho_{D:ATP}
  \label{eq:d-on-wall}
\end{multline}
\begin{multline}
  \frac{\partial \sigma_{DE}}{\partial t} = -k_{de}\sigma_{DE}+k_E\sigma_D\rho_E\hspace{3cm}
  \label{eq:FifthPDE}
\end{multline}

where $\rho$ is cytoplasmic protein density ($\mu m^{-3}$), $\sigma$
is membrane bound density ($\mu m^{-2}$), $\mathcal{D}_D$ and
$\mathcal{D}_{E}$ th diffusions constants for MinD and MinE,
respectively, $k_D^{\textrm{ADP $\rightarrow$ ATP}}$ the rate of
conversion from MinD:ADP to the MinD:ATP complex, $k_D$ the rate of
MinD:ATP attachement to the membrane when no protein is already
attached there, $k_{dD}$ the increase of this rate when MinD:ATP is
present on the membrane, $k_{de}$ the rate of hydrolisis of the
MinD:MinE:ATP complex, $k_E$ the rate of cytoplasmic MinE binding to
membrane bound MinD:ATP complex, and $d_w$ is the distance from the
point in space to the closest wall (which will always be
perpendicular to this distance).  $\delta(d_w)$ has units of
$\mu m^{-1}$ and will be zero everywhere except at the wall.
Equations \ref{eq:d-on-wall} and \ref{eq:FifthPDE} will only be
relevent at the membrane because the membrane bound density values
will have no meaning in the cytoplasm.

@@

Huang's simulations use total MinD and MinE concentrations of
$1,000/\mu m$ and $350/\mu m$, respectively, in a cylindrical cell of
radius $0.5\mu m$, and in our (non-cylindrical) cells we use the same
number of proteins per unit volume.  These concentration values are
1273 $\mu m^{-3}$ and 446 $\mu m^{-3}$, respectively. \fixme{It's been
  a while so might want to check this to make sure it's true.}  A 3D
grid is then constructed in cartesian coordinates, with a grid spacing
of .05\micron.

@@

We mean to investigate the geometric limits of the Min system
oscillations as observed by Mannik \emph{et
  al.}~\cite{mannik2012robustness}, so we model the Min system in a
variety of cell shapes and sizes.  Here we present a selection of
these, first naturally occuring pill-shaped cells and then a number of
flattened out shapes which reflect the experiments of Mannik \emph{et
  al.}, in which bacteria are confined within a thin slit of height
$0.25\micron$ ~\cite{mannik2012robustness}. Viewed from the top down
the cells will show the shapes described below and viewed from the
side they show at their edges a semicircular protrusion (one may
imagine the shape of a pancake). We have simulated and analyzed a
number of these flattened shapes in addition to those shown below,
including equilateral and iscosolese triangle shapes, shapes shaped
like greater than signs, '$>$', and four-pronged 'starfish' shapes.
For all of these we have varied the cell size, expanding the two
dimensions that are tangent to the flattened surface from very small,
'steady state' scales (which we discuss below) to scales twice that of
the published Mannik shapes.  Our conclusions are based on all of this
data, but we present here specific, illustrative examples: two
flattened shapes that are designed to replicate those published by
Mannik and two `stadium' shapes that respectively have the same aspect
ration, thickness, and volume as the two Mannik shapes.  Viewed from
the top down, the stadium shapes appear as rectangles with
semi-circular end caps on their long axis ends.

@@

\fixme{Is the following Kubitscheck paragraph needed?}
Kubitschek has shown in multiple experiments that at the time of cell
division cells have a volume that is within a range of roughly
$1\micron^3$ to $2\micron^3$~\cite{kubitschek1990cell,
  kubitschek1968linear}.  We follow Huang's
simulations\cite{huang2003dynamic} and Mannik's experiments and model
cells that are slightly larger than this range.

\begin{figure*}
  %\includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-exact/plots/image-plot}
  \includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-exact/plots/single-image-plot}
  \includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-full_array/plots/single-image-plot}
  \caption{Contour plot images of the concentration of each protein
    species in a natural pill-shaped bacterium at regular intervals in
    time (one second intervals), with darker regions indicating higher
    concentration. The upper plots shows results from the
    deterministic model and the lower shows results from the
    stochastic model.  The cells pictured are $4\micron$ in
    length, measured from end to end.  The order of frames is such
    that individual MinD proteins begin at the bottom of the plot (in
    the MinD:ATP state in the cytoplasm), and progress upward until
    they reach the MinE:MinD:ATP membrane-bound complex.  At that
    point, they will spontaneously dissociate into cytoplasmic MinE
    (the top row) and the starting state of cytoplasmic MinD:ADP.  The
    densities plotted are integrated along the axis orthogonal to the
    page, and the color scale is chosen separately for each species
    with black as the maximum value over space and time.  The
    stochastic densities are smeared out over space using a guassians
    approximation developed by Zhang et all~\cite{zhang2007gaussian}}.
  \label{image-p}
\end{figure*}

Huang \emph{et al}~\cite{huang2003dynamic} has performed a linear
stability analysis on the model which shows an upper limit on a steady
state solution of a $2\mu m$ wavelength.  Cells with dimensions longer
than this are unstable and show oscillatory behavoir in this
direction, while cells that are shorter relax into a motionless steady
state.  We have similarly performed a stability analysis in an
infinite slab with a thickness equal to that of our flattened cells'
($0.25\mu m$) and have found that an equivalent stability limit for
half wavelengths is $2.13\mu m$. As expected, when decreasing the
lengths and widths of our simulated flattened cells so that the
longest distance across the cell is less than this length, the cells
stop exhibiting any oscillalitory behavior.  The deterministic model
relaxes into a motionless state and the stochastic model exhibits
multiple random locations of small maximization.

@@

This simple cell shape is a good starting point for observing in
detail the dynamic interaction between the different stages of protein
that lead to their qualitatively distinct behavoir. Figure
\ref{image-p} shows the process in a series of frames taken from
simulation animations for both the deterministic and stochastic
models.  We have 'smeared out' the stochastic concentrations in a
manner meant to reproduce the images shown by diffraction limited
flourescence microscopy.  We do so using the two dimensional gaussian
approximation developed by Zhang et all~\cite{zhang2007gaussian}.  In
this approximation we use a numerical aperture value of 1.3, which is
the same as used by mannik experimentally, and a wavelength of
$650nm$.  Each frame is 2.5 seconds ahead of the last, and each image
shows the concentration of a given state of protein (of the five
described in the reaction model) summed over the coordinate orthogonal
to the page.  The color scale for each protein state is set according
to the maximum values observed.

Figure~\ref{image-p} begins about 300 seconds into the simulation and
shows a typical period of oscillation that in the pill shape repeates
itself with a high degree of regularity, for both the deterministic
and stochastic models.  At $t=0$ the cytoplasmic MinE is
concentrated in the upper portion of the cell and is diffusing
downward, where it will react with and stick to wall-bound MinD that
is concentrated on the membrane in the lower portion of the cell.
This process is in fact already well on its way, as seen by the
significant concentration of wall bound MinD:MinE which is creeping
down to the lower corner, removing the MinD from the membrane as it
goes.  The removed MinD (cytoplasmic MinD:ADP) diffuses freely for a
time before spontenously undergowing nucleotide exchange and shifting
back into the MinD:ATP that is able to attach to the membrane.
Diffusing a small distance away from the bottom of the cell will bring
the protein into the MinE collection, where reactions will prevent it
from collecting for any significant time on the membrane.  Diffusing
further towards the upper end, however, while the MinE has not yet
collected there, will allow MinD:ATP to accumulate with the other
MinD:ATP, which is seen at seconds 10 through 15 in
Figure~\ref{image-p}.  Eventually the MinE creeps all the way down to
the bottom of the cell where it slowly runs out of wall bound MinD to
react with, before diffusing back up and starting again on the upper
portion of the cell. This is seen in seconds 15 through 20 in the
figure.  \fixme{There is another paragraph after this in the latex
  that I commented out because I thought we might be spending too much
  time on the description}

@@

We test cells of radius $.5\micron$ and of lengths $2\micron$,
$3\micron$, $4\micron$, and $5\micron$ measured from the tip of one
end cap to the tip of the other \fixme{I'm running the 2,3,5 pills
  again to make sure that everything in this paragraph is true}. In
each case, we observe in the deterministic model a quick
establishement (within a few periods) of regular oscillatory behavoir
that lasts indefinitely.  This temporal regularity can be seen in
Figure~\ref{corr-pill}, which shows the temporal correlation of the
total MinD protein located in the the two polar sections of the cell.
In the case of the stochastic model the pills exhibit coherence times
of over 7 periods and in the case of the determinisitic model cohere
indefinately.

@@

While the stochastic results show periods that seem to have a high
degree of regularity as well as similarity to the deterministic
results, they do differ in the precise locations of the maxima.  While
the deterministic model shows maxima that are every time
concentrated in the center of one of the cell ends, with a symmetric
smear of protein going out laterally to either side, the stochastic
model shows maxima that, while always located somehwere in the polar
region of the cell, occur at different positions within this region.
This is reflected in Figure~\ref{image-p}, which shows periods that
are the same and qualitative behavoir that is approximately the same,
but also that there is variation in the details of the stochastic
model maxima location.

\fixme{Is the table of the cell periods needed? I'm not sure if we
  have a experiment to compare this to} The stochastic results show
oscillitions that are similarly regular in time, with periods that
nearly match those of the deterministic (they differ roughly by a
factor of .05) \fixme{does this square with the correlation data
  periods?}. The regularity of the pill shape cell periods is
discussed below, in our analysis of time correlation data \fixme{not
  written yet}.  These periods increase with cell length, as shown in
Table~\ref{tab:pill-periods}, in agreement with the results of Huang
\emph{et al.}.

\begin{table}
  \begin{tabular}{|r|c|c|c|l|}
    \hline Length($\mu$) & 2.50 & 3.00 & 4.00 & 5.00\\ \hline
    Period(sec) & sim & 33 & 38 & 48 \\ \cline{2-2} \hline
  \end{tabular}
  \caption{Period of oscillation according to length of cell for
    pill-shaped cells.}\label{tab:pill-periods}
\end{table}

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{../paper/plot-ave}
  \caption{We display here arrows depicting successive maxima in space
    and and time overlayed on a color plot of the total MinD density
    averaged over the same time period.  The simulation time covered
    for the flattened cells is \fixme{340} seconds, which is the same
    period of time depicted in the experimental data plots of Mannik
    \emph{et al.}~\cite{mannik2012robustness}.  For the wild type pill
    shape, we only cover \fixme{???} seconds, in order to provide a
    useful comparison due to its shorter oscillation period.  The top
    row shows plots published by Mannik of the MinD maxima behavoir
    and the bottom two rows show our simulations using the stochastic
    and deterministic models, respectively.  We simulated
    approximations to the two shapes observed by Mannik, which we call
    \emph{shape A} and \emph{shape B}.  In addition, we studied two
    flattened stadium shapes which we call \emph{stadium A} and
    \emph{stadium B} corresponding in aspect ratio and thickness to
    the two experimental shapes.  The spatial length scale of all the
    figures shown was identical. Each of the flattened cell shapes
    uses the same color scale for the number of proteins per unit
    area.  Finally, we display the natural pill shape, which was also
    featured in Figs.~\ref{image-p} and~\ref{corr-pill}, with a
    different color scale to reflect the thicker cell containing more
    proteins per cross-sectional area.  }
  \label{randst-plot-ave}
\end{figure*}

@@

As describe in Section~\ref{sec:model-method-shapes}, we will focus on
simulation of four illustrative flattened cell shapes: two shapes
created to replicate those published by
Mannik~\cite{mannik2012robustness} (shape~\emph{A} and
shape~\emph{B}), and for comparison two symmetrical stadium shapes
that have the identical aspect ratio, thickness and volume (stadium
\emph{A} and stadium \emph{B}).  These can be seen in
Figure~\ref{randst-plot-ave}.  These stadium shapes enable us to
distinguish between the effect of cellular shape irregularity and
assymetry, and that of overall size and aspect ratio.  We note that
shape \emph{B} is more irregular than shape \emph{A}, and in
particular features a prominant region of concavity on its left-hand
side.  For each of these shapes (in addition to the wild type pill
shape discussed previously) we have simulated in excess of \fixme{???}
seconds of evolution of the MinD system using the deterministic
version of Huang's model~\cite{huang2003dynamic}, as well as the
stochastic model.

From each of these flattened simulations, we have chosen a typical 350
second segment to compare with the results of Mannik exemplified in
arrow plots in which the arrow heads show the location of sequential
MinD maxima within the cell (over the same time
interval)~\cite{mannik2012robustness}.  This comparison is shown in
Fig.~\ref{randst-plot-ave}.  In addition to arrows between successive
maxima in space and time, we plot as a colored background the density
of MinD protiens averaged over the same time period.  We note that we
have manually verified that our (computer-generated) arrow plots also
reflect a human interpretation of a movie of the same data.  Finally,
for comparison we present the same plot for a wild-type cell, with a
time period of \fixme{???}  seconds to account for its short period of
oscillation.

>>>>

However, Mannik \emph{et al.} have also shown that there are
limitations to this robust capability to
oscillate~\cite{mannik2010bacteria,mannik2009bacterial}. They have
experimentally forced E. Coli cells into silicon microfabricated
channels of $0.25\micron$ thickness. Upon entering the channels, the
cells undergo a rapid mechanical deformation in which they widen along
their short axis parallel to the channel plane, followed by a slower,
broadening out along both free axes.  This deformation results in very
wide (they can reach widths of over $5\micron$), flattened cells that
when view from the top down show very irregular and assymetric shapes.
While these cells are still able to divide into surprisingly equal
volumes, the MinD oscillations within lose their regularity, both
temporally and spatially. Seen with flourescent microscopy (after
using common genetic techniques to allow the cells to flouresce), the
MinD seem to maximize in multiple locations within the cell, in a
seemingly random sequence. This allows for an opportunity to test
previously studied MinD simulation models against more extreme
experimental cases than have been seen thus far. \fixme{Perhaps this
  sentence should come later after we've introduced the simulation
  research that's been done.}

@@

\begin{multline}
  \frac{\partial \rho_{D:ADP}}{\partial t} = \mathcal{D}_D\nabla^2\rho_{D:ADP}-k_D^{ADP\rightarrow ATP}\rho_{D:ADP}\\
  +\delta(d_w)k_{de}\sigma_{DE},\hspace{3.4cm}
\end{multline}
\begin{multline}
  \frac{\partial \rho_{D:ATP}}{\partial t} = \mathcal{D}_D\nabla^2\rho_{D:ATP}+k_D^{ADP\rightarrow ATP}\rho_{D:ADP}\\
  -\delta(d_w)[k_D+k_{dD}(\sigma_D+\sigma_{DE})]\rho_{D:ATP}
\end{multline}
\begin{multline}
  \frac{\partial \rho_E}{\partial t} = \mathcal{D}_E\nabla^2\rho_E+\delta(d_w)k_{de}\sigma_{DE}
  -\delta(d_w)k_E \sigma_D \rho_E
\end{multline}
\begin{multline}
  \frac{\partial \sigma_D}{\partial t} = -k_E\sigma_D\rho_E
  +[k_D+k_{dD}(\sigma_D+\sigma_{DE})]\rho_{D:ATP}
  \label{eq:d-on-wall}
\end{multline}
\begin{multline}
  \frac{\partial \sigma_{DE}}{\partial t} = -k_{de}\sigma_{DE}+k_E\sigma_D\rho_E\hspace{3cm}
  \label{eq:FifthPDE}
\end{multline}
where $\rho$ is cytoplasmic protein density ($\micron^{-3}$), $\sigma$
is membrane bound density ($\micron^{-2}$), $\mathcal{D}_D$ and
$\mathcal{D}_{E}$ th diffusions constants for MinD and MinE,
respectively, $k_D^{\textrm{ADP $\rightarrow$ ATP}}$ the rate of
conversion from MinD:ADP to the MinD:ATP complex, $k_D$ the rate of
MinD:ATP attachement to the membrane when no protein is already
attached there, $k_{dD}$ the increase of this rate when MinD:ATP is
present on the membrane, $k_{de}$ the rate of hydrolisis of the
MinD:MinE:ATP complex, $k_E$ the rate of cytoplasmic MinE binding to
membrane bound MinD:ATP complex, and $d_w$ is the distance from the
point in space to the closest wall.  The Dirac delta function
$\delta(d_w)$ has units of $\micron^{-1}$ and is zero everywhere except
at the wall.  Equations \ref{eq:d-on-wall} and \ref{eq:FifthPDE} are
only relevent at the membrane because the membrane-bound density
values have no meaning in the cytoplasm.

@@

Huang's simulations use total MinD and MinE concentrations of
$1,000/\micron$ and $350/\micron$, respectively, in a cylindrical cell of
radius $0.5\micron$, and in our (non-cylindrical) cells we use the same
number of proteins per unit volume.  These concentration values are
$1273\micron^{-3}$ and $446\micron^{-3}$, respectively. \fixme{It's been
  a while so might want to check this to make sure it's true.}  We
construct a 3D grid in cartesian coordinates, with a grid spacing of
.05\micron.

@@

We mean to investigate the geometric limits of the Min system
oscillations as observed by Mannik \emph{et
  al.}~\cite{mannik2012robustness}, so we model the Min system in
several cell shapes and sizes.  Here we present a selection of these,
beginning with naturally occuring pill-shaped cells, followed by a
number of flattened out shapes which reflect the experiments of Mannik
\emph{et al.}, in which bacteria are confined within a thin slit of
height $0.25\micron$ ~\cite{mannik2012robustness}.  Viewed from the
top down the cells have the shapes described below and viewed from the
side they have at their edges a semicircular protrusion (one may
imagine the shape of a pancake).
%
In this paper we focus on four specific flattened cell shapes.  Two of
these shapes replicate those published by Mannik, and the other two
are `stadium' shapes that respectively have the same aspect ratio,
thickness, and volume as the two Mannik shapes.  Viewed from the top
down, these stadium shapes appear as rectangles with semi-circular end
caps on the long axis ends.

@@

%% \fixme{Is the following Kubitscheck paragraph needed?}
%% Kubitschek has shown in multiple experiments that at the time of cell
%% division cells have a volume that is within a range of roughly
%% $1\micron^3$ to $2\micron^3$~\cite{kubitschek1990cell,
%%   kubitschek1968linear}.  We follow Huang's
%% simulations\cite{huang2003dynamic} and Mannik's experiments and model
%% cells that are slightly larger than this range.

\begin{figure*}
  %\includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-exact/plots/image-plot}
  \begin{center}
    \includegraphics[width=0.8\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-exact/plots/single-image-plot}\\
    \vspace{-2.5em}
    \includegraphics[width=0.8\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-full_array/plots/single-image-plot}
    \vspace{-3em}
  \end{center}
  \caption{Contour plot images of the concentration of each protein
    species in a natural pill-shaped bacterium at regular intervals in
    time (one second intervals), with darker regions indicating higher
    concentration. The upper plots shows results from the
    deterministic model and the lower shows results from the
    stochastic model.  The cells pictured are $4\micron$ in length,
    measured from end to end.  The order of frames is such that
    individual MinD proteins begin at the bottom of the plot (in the
    MinD:ATP state in the cytoplasm), and progress upward until they
    reach the MinE:MinD:ATP membrane-bound complex.  At that point,
    they will spontaneously dissociate into cytoplasmic MinE (the top
    row) and the starting state of cytoplasmic MinD:ADP.  The
    densities plotted are integrated along the axis orthogonal to the
    page.  The stochastic densities are broadened using the gaussian
    smearing approach of Zhang \emph{et
      al.}~\cite{zhang2007gaussian}}.
  \label{image-p}
\end{figure*}

Huang \emph{et al}~\cite{huang2003dynamic} has performed a linear
stability analysis on the model which shows an upper limit on a steady
state solution of a $2\micron$ wavelength.  Cells with dimensions longer
than this show spontaneous oscillatory behavoir in this
direction, while cells that are shorter relax into a motionless steady
state.  We have similarly performed a stability analysis in an
infinite slab with a thickness equal to that of our flattened cells
($0.25\micron$) and have found that an equivalent stability limit for
half wavelength is $2.13\micron$. As expected, when decreasing the
lengths and widths of our simulated flattened cells so that the
longest distance across the cell is less than this length, the cells
stop exhibiting any oscillatory behavior.  The deterministic model
relaxes into a motionless state and the stochastic model exhibits
random fluctuations without spatial oscillations.

\fixme{WHERE TO SAY??? : I think we should describe the ``starry
  night'' effect in the stochastic model.}

@@

This simple cell shape is a good starting point for observing in
detail the dynamic interaction between the different stages of protein
that lead to their qualitatively distinct behavoir. Figure
\ref{image-p} shows the process in a series of frames taken from
simulation animations for both the deterministic and stochastic
models.  We have 'smeared out' the stochastic concentrations in a
manner meant to reproduce the images shown by diffraction limited
flourescence microscopy.  We do so using the two dimensional gaussian
approximation developed by Zhang \emph{et al.}~\cite{zhang2007gaussian}.  In
this approximation we use a numerical aperture value of 1.3, which is
the same as used by Mannik experimentally, and a wavelength of
$650nm$.  Each frame is 2.5 seconds ahead of the last, and each image
shows the concentration of a given state of protein (of the five
described in the reaction model) summed over the coordinate orthogonal
to the page.

Figure~\ref{image-p} begins about 300 seconds into the simulation and
shows one period of oscillation.  At $t=0$ the cytoplasmic MinE is
concentrated in the upper portion of the cell and is diffusing
downward, where it will react with and stick to wall-bound MinD that
is concentrated on the membrane in the lower portion of the cell.
This process is in fact already well on its way, as seen by the
significant concentration of wall bound MinD:MinE which is creeping
down to the lower corner, removing the MinD from the membrane as it
goes.  The removed MinD (cytoplasmic MinD:ADP) diffuses freely for a
time before spontenously undergoing nucleotide exchange and shifting
back into the MinD:ATP that is able to attach to the membrane.
Diffusing a small distance away from the bottom of the cell will bring
the protein into the MinE \fixme{collection ??}, where reactions will prevent it
from collecting for any significant time on the membrane.  Diffusing
further towards the upper end, however, while the MinE has not yet
collected there, will allow MinD:ATP to accumulate with the other
MinD:ATP, which is seen at seconds 10 through 15 in
Figure~\ref{image-p}.  Eventually the MinE creeps all the way down to
the bottom of the cell where it slowly runs out of wall bound MinD to
react with, before diffusing back up and starting again on the upper
portion of the cell. This is seen in seconds 15 through 20 in the
figure.  \fixme{There is another paragraph after this in the latex
  that I commented out because I thought we might be spending too much
  time on the description}

@@

@@

\fixme{add explanation of correlation function somewhere in the
  following paragraph.  The math could go in theory/methods section.}
{
  \color{red}
  \begin{equation}
    C(\tau) = \frac{\int
      (N_{\textit{top}}(t) - \bar N_{\textit{top}})
      (N_{\textit{bottom}}(t+\tau) - \bar N_{\textit{bottom}})dt
    }{
      \int
      (N_{\textit{top}}(t) - \bar N_{\textit{top}})
      (N_{\textit{bottom}}(t) - \bar N_{\textit{bottom}})dt
    }
  \end{equation}
  \begin{equation}
    C(\tau) \propto \int
      (N_{\textit{top}}(t) - \bar N_{\textit{top}})
      (N_{\textit{bottom}}(t+\tau) - \bar N_{\textit{bottom}})dt
  \end{equation}
  where...
  \begin{equation}
    C(\tau) = -\cos\left(\frac{2\pi\tau}{T}\right) e^{-\frac{\tau}{\tau_0}}
  \end{equation}
  where $T$ is the period and $\tau_0$ is the coherence time.
}

We present simulations of a cell with radius $.5\micron$ and length
$4\micron$, measured from the tip of one end cap to the tip of the
other. We observe in the deterministic model a quick establishement
(within a few periods) of regular oscillatory behavoir that lasts
indefinitely.  This temporal regularity can be seen in
Figure~\ref{corr-pill}, which shows the temporal correlation of the
total MinD protein located in the the two polar sections of the cell.
In the case of the stochastic model the pill exhibits coherence times
of over 7 periods and in the case of the determinisitic model
oscillate indefinately with complete coherence.

While the stochastic results show periods that seem to have a high
degree of regularity as well as similarity to the deterministic
results, they do differ in the precise locations of the maxima.  While
the deterministic model shows maxima that are every time
concentrated in the center of one of the cell ends, with a symmetric
smear of protein going out laterally to either side, the stochastic
model shows maxima that, while always located somehwere in the polar
region of the cell, occur at different positions within this region.
This is reflected in Figure~\ref{image-p}, which shows periods that
are the same and qualitative behavoir that is approximately the same,
but also that there is variation in the details of the stochastic
model maxima location.  \fixme{I think we should discuss stochastic
  picture earlier.}

\begin{figure*}
  \centering
  \includegraphics[width=0.8\textwidth]{../paper/plot-ave}
  \caption{We display here arrows depicting successive maxima in space
    and and time overlayed on a color plot of the total MinD density
    averaged over the same time period.  The simulation time covered
    for the flattened cells is \fixme{340} seconds, which is the same
    period of time depicted in the experimental data plots of Mannik
    \emph{et al.}~\cite{mannik2012robustness}.  For the wild type pill
    shape, we only cover \fixme{???} seconds, in order to provide a
    useful comparison due to its shorter oscillation period.  The top
    row shows plots published by Mannik of the MinD maxima behavoir
    and the bottom two rows show our simulations using the stochastic
    and deterministic models, respectively.  We simulated
    approximations to the two shapes observed by Mannik, which we call
    \emph{shape A} and \emph{shape B}.  In addition, we studied two
    flattened stadium shapes which we call \emph{stadium A} and
    \emph{stadium B} corresponding in aspect ratio and thickness to
    the two experimental shapes.  The spatial length scale of all the
    figures shown was identical. Each of the flattened cell shapes
    uses the same color scale for the number of proteins per unit
    area.  Finally, we display the natural pill shape, which was also
    featured in Figs.~\ref{image-p} and~\ref{corr-pill}, with a
    different color scale to reflect the thicker cell containing more
    proteins per cross-sectional area.  }
  \label{randst-plot-ave}
\end{figure*}

@@

As describe in Section~\ref{sec:model-method-shapes}, we will focus on
simulation of four illustrative flattened cell shapes: two shapes
created to replicate those published by
Mannik~\cite{mannik2012robustness} (shape~\emph{A} and
shape~\emph{B}), and for comparison two symmetrical stadium shapes
that have the identical aspect ratio, thickness and volume (stadium
\emph{A} and stadium \emph{B}), shown in Figure~\ref{randst-plot-ave}.
These stadium shapes enable us to distinguish between the effect of
cellular shape irregularity and assymetry, and that of overall size
and aspect ratio.  We note that shape \emph{B} is more irregular than
shape \emph{A}, and in particular features a prominant region of
concavity on its left-hand side.  For each of these shapes (in
addition to the wild type pill shape discussed previously) we have
simulated in excess of \fixme{???}  seconds of evolution of the MinD
system using the deterministic version of Huang's
model~\cite{huang2003dynamic}, as well as the stochastic model.

From each of these flattened simulations, we have chosen a typical 350
second segment to compare with the results of Mannik exemplified in
arrow plots in which the arrow heads show the location of sequential
MinD maxima within the cell (over the same time
interval)~\cite{mannik2012robustness}.  In addition to arrows between
successive maxima in space and time, we plot as a colored background
the density of MinD protiens averaged over the same time period.  We
note that we have manually verified that our (computer-generated)
arrow plots also reflect a human interpretation of a movie of the same
data.  Finally, for comparison we present the same plot for a
wild-type cell, with a time period of \fixme{???}  seconds to account
for its short period of oscillation.
